timesliced_random_cell_selection

### 2018-10-31 src/raster_sjl.Rmd
finding lon lat wgs84

We can plot the different layers inside the raster brick.

Plot the top layer
```{r}
plot(rasbr[[1]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
```

plot layer 20
```{r}
plot(rasbr[[20]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
```

@@

The next piece of data we need is the observation data itself. I've sent you the observations.csv in the email with this code.
```{r}
obs <- read.csv("../data/bio/observations.csv") #change this path to your .csv location
head(obs)
```

>>>>

We can plot the different layers (which represent depth) inside the raster brick.

Plot the top layer
```{r}
plot(rasbr[[1]]) #the double brackets tell R which layer to plot
```

plot layer 20
```{r}
plot(rasbr[[20]]) 
```

@@

The next piece of data we need is the observation data itself. I've sent you the observations.csv in the email with this code.
```{r}
obs <- read.csv("../data/bio/observations.csv") #change this path to your .csv location
head(obs)
```
A note about obs$depth. This is not a measurement, but relates to the layer in the raster bricks. E.G. the observation in row one was found at rasbr[[15]], the observation in row 3 at rasbr[[17]]. I created this field myself.

### 2018-10-31 src/timesliced_random_cell_selection.Rmd
finding lon lat wgs84

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("raster", "dplyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

you have a unique_cell raster already (glo_unique_cell) (output/env/glo_unique_cell.tif)

@@

```{r}
rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4)
  randtemp$month <- substr(rand_list[i], 6, 7)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

ok now you can start to populate the rest of the background point information. The background points file should mirror as much as possible the occurence file

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("raster", "dplyr", "data.table", "sp")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

# the show
you have a unique_cell raster already (glo_unique_cell) (output/env/glo_unique_cell.tif)

@@

```{r}
rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

# populating the background points data with... data

ok now you can start to populate the rest of the background point information. The background points file should mirror as much as possible the occurence file. Maybe it would be best to merge the files together, and them split them again later? would save looping everything to repeat the process through multiple files...

```{r}
library(data.table)
filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
background_all <- rbindlist(lapply(filenames,fread))
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
```

get the columns from the occurence data so you know what to add
```{r}
occ_column <- colnames(data_aea)
occ_column
```

ok so have cell_id, year, month, decimalLongitude.1(meters - in background_all its longitude), decimalLatitude.1 (meters - in background_all its latitude), depth_layer_no.

Easy first one - add an id based on id_depth
```{r}
background_all$id <- background_all$id_depth
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
```

ok lets's get the real decimalLatitude and decimalLongitude.
Will need to plot all point, extract the lonlat coordinates

1) load the file as a spatialdatapointframe. give it the aea coordinate

```{r}
library(sp)
xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
head(sp_background)
```
```{r}
plot(sp_background)
```

2) reproject to a wgs

```{r}
sp_background <- spTransform(sp_background, CRS = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
plot(sp_background)
```
3) extract lonlat, add to dataframe

```{r}
head(sp_background)
```

save as csv
